Analysis for Biracial Grade Retention Project

### 2022-09-03 analysis/analysis.Rmd
Change all models to use survey design effects

```{r build-models, results='asis'}
model0 <- glm(below_exp_grade~race, data=acs, family=binomial)
model1 <- update(model0, 
                 .~.+I(year-2010)*current_grade+state+metro+sex)
model2 <- update(model1, 
                 .~.+foreign_born+foreign_born_mother+foreign_born_father+
                   spk_eng+spk_eng_mother+spk_eng_father)
model3 <- update(model2, .~.+degree_mother+degree_father)
model4 <- update(model3, .~.+sqrt(family_income/10000)+own_home+parents_married)

@@

```{r monoracial-results, warning=FALSE, results='asis'}
results_monoracial <- map(list(model0, model1, model2, model3, model4),
                          function(x) {
                            marg(x, "race", type="effects", 
                                 vcov_mat=vcovCL(x, cluster=acs$cluster),
                                 at_var_interest=c("White","Black","Indigenous",
                                                   "Asian","Latino"))[[1]][-1,]
                          })

@@

```{r models-by-grade, warning=FALSE, fig.height=12, fig.cap="Probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level. Panels compare results when data are restricted to certain grade ranges."}
#run separate models by grade groups
model4_elem <- update(model4, data=subset(acs, current_grade<="5th"))
model4_middle <- update(model4, data=subset(acs, current_grade>"5th" & 
                                        current_grade<"9th"))
model4_hs <- update(model4, data=subset(acs, current_grade>="9th"))

coef_table_elem <- calculate_marg_means(model4_elem, 
                                        cluster_var=subset(acs, current_grade<="5th")$cluster)
coef_table_middle <- calculate_marg_means(model4_middle,
                                          cluster_var=subset(acs, current_grade>"5th" & 
                                        current_grade<"9th")$cluster)
coef_table_hs <- calculate_marg_means(model4_hs,
                                      cluster_var=subset(acs, current_grade>="9th")$cluster)

@@

```{r models-parent-link, warning=FALSE, fig.height=12, fig.cap="Probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level. Panels compare results depending on whether results are restricted only to children linked to a head of household or not."}
#run separate models by grade groups
model4_best_link <- update(model4, data=subset(acs, momrule==11 & poprule==11))

coef_table_best_link <- calculate_marg_means(model4_best_link, 
                                        cluster_var=subset(acs, momrule==11 &
                                                             poprule==11)$cluster)

@@

```{r model-dummied}
model_full <- glm(below_exp_grade~
                    year*(current_grade_1st+current_grade_2nd+current_grade_3rd+
                            current_grade_4th+current_grade_5th+
                            current_grade_6th+current_grade_7th+
                            current_grade_8th+current_grade_9th+
                            current_grade_10th+current_grade_11th+
                            current_grade_12th)+
                    race_black+race_indigenous+race_asian+race_latino+
                    race_black_white+race_black_indigenous+race_black_latino+
                    race_black_asian+race_white_indigenous+race_white_latino+
                    race_white_asian+race_indigenous_latino+
                    race_indigenous_asian+race_latino_asian+
                    sex_female+
                    metro_non_metro+metro_central_city+metro_metro_non_central+
                    metro_metro_indeterminable+
                    state_alaska+state_arizona+state_arkansas+state_california+
                    state_colorado+state_connecticut+state_dc+state_delaware+
                    state_florida+state_georgia+state_hawaii+state_idaho+
                    state_illinois+state_indiana+state_iowa+state_kansas+
                    state_kentucky+state_louisiana+state_maine+state_maryland+
                    state_massachusetts+state_michigan+state_minnesota+
                    state_mississippi+state_missouri+state_montana+
                    state_nebraska+state_nevada+state_new_hampshire+
                    state_new_jersey+state_new_mexico+state_new_york+
                    state_north_carolina+state_north_dakota+state_ohio+
                    state_oklahoma+state_oregon+state_pennsylvania+
                    state_rhode_island+state_south_carolina+state_south_dakota+
                    state_tennessee+state_texas+state_utah+state_vermont+
                    state_virginia+state_washington+state_west_virginia+
                    state_wisconsin+state_wyoming+
                    +sqrt(family_income/10000)+parents_married+own_home+
                    degree_mother+degree_father+
                    spk_eng+spk_eng_mother+spk_eng_father+
                    foreign_born_mother+foreign_born_father+foreign_born, 
                  data=acs_dummied, family=binomial)
```

>>>>

```{r create-design}
# strata make the models take forever, but do not seem to change anything
# so I will leave them out
acs_design <- svydesign(ids=~cluster, weights=~perwt, data=acs)
```

```{r build-models, results='asis'}
model0 <- svyglm(below_exp_grade~race, design=acs_design, family=binomial)
model1 <- update(model0, 
                 .~.+I(year-2010)*current_grade+state+metro+sex)
model2 <- update(model1, 
                 .~.+foreign_born+foreign_born_mother+foreign_born_father+
                   spk_eng+spk_eng_mother+spk_eng_father)
model3 <- update(model2, .~.+degree_mother+degree_father)
model4 <- update(model3, .~.+sqrt(family_income/10000)+own_home+parents_married)

@@

```{r monoracial-results, warning=FALSE, results='asis'}
results_monoracial <- map(list(model0, model1, model2, model3, model4),
                          function(x) {
                            marg(x, "race", type="effects", 
                                 at_var_interest=c("White","Black","Indigenous",
                                                   "Asian","Latino"))[[1]][-1,]
                          })

@@

```{r models-by-grade, warning=FALSE, fig.height=12, fig.cap="Probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level. Panels compare results when data are restricted to certain grade ranges."}
#run separate models by grade groups
model4_elem <- update(model4, design=subset(acs_design, current_grade<="5th"))
model4_middle <- update(model4, design=subset(acs_design, current_grade>"5th" & 
                                        current_grade<"9th"))
model4_hs <- update(model4, design=subset(acs_design, current_grade>="9th"))

coef_table_elem <- calculate_marg_means(model4_elem)
coef_table_middle <- calculate_marg_means(model4_middle)
coef_table_hs <- calculate_marg_means(model4_hs)

@@

```{r models-parent-link, warning=FALSE, fig.height=12, fig.cap="Probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level. Panels compare results depending on whether results are restricted only to children linked to a head of household or not."}
#run separate models by grade groups
model4_best_link <- update(model4, design=subset(acs_design, momrule==11 & poprule==11))

coef_table_best_link <- calculate_marg_means(model4_best_link)

@@

```{r model-dummied}
acs_dummied_design <- svydesign(ids=~cluster, weights=~perwt, data=acs_dummied)

model_full <- svyglm(below_exp_grade~
                       year*(current_grade_1st+current_grade_2nd+current_grade_3rd+
                               current_grade_4th+current_grade_5th+
                               current_grade_6th+current_grade_7th+
                               current_grade_8th+current_grade_9th+
                               current_grade_10th+current_grade_11th+
                               current_grade_12th)+
                       race_black+race_indigenous+race_asian+race_latino+
                       race_black_white+race_black_indigenous+race_black_latino+
                       race_black_asian+race_white_indigenous+race_white_latino+
                       race_white_asian+race_indigenous_latino+
                       race_indigenous_asian+race_latino_asian+
                       sex_female+
                       metro_non_metro+metro_central_city+metro_metro_non_central+
                       metro_metro_indeterminable+
                       state_alaska+state_arizona+state_arkansas+state_california+
                       state_colorado+state_connecticut+state_dc+state_delaware+
                       state_florida+state_georgia+state_hawaii+state_idaho+
                       state_illinois+state_indiana+state_iowa+state_kansas+
                       state_kentucky+state_louisiana+state_maine+state_maryland+
                       state_massachusetts+state_michigan+state_minnesota+
                       state_mississippi+state_missouri+state_montana+
                       state_nebraska+state_nevada+state_new_hampshire+
                       state_new_jersey+state_new_mexico+state_new_york+
                       state_north_carolina+state_north_dakota+state_ohio+
                       state_oklahoma+state_oregon+state_pennsylvania+
                       state_rhode_island+state_south_carolina+state_south_dakota+
                       state_tennessee+state_texas+state_utah+state_vermont+
                       state_virginia+state_washington+state_west_virginia+
                       state_wisconsin+state_wyoming+
                       +sqrt(family_income/10000)+parents_married+own_home+
                       degree_mother+degree_father+
                       spk_eng+spk_eng_mother+spk_eng_father+
                       foreign_born_mother+foreign_born_father+foreign_born, 
                     design=acs_dummied_design, family=binomial)
```
